---
title: "DSBikeshare"
author: "Jason Phelps"
date: "10/14/2020"
output: html_document
runtime: shiny
---

---
title: "Clarkson Class Data"
author: "Jason Phelps"
date: "16 Sep 2020"
output: html_notebook
---

The intention for this file is to conduct analysis on clarkson's enrollment data.
## Load packages
```{r}
library(tidyverse)
library(janitor)
library(readxl)
library(skimr)
library(summarytools)
library(tidyr)
library(lubridate)

```
Read .csv file and add in variables through 'r' not directly in excel

```{r}
dfa= read_excel("DS_bikeshare.xlsx") %>%
clean_names()
  
```

creating a second data frame to eliminate any useless entries into the data set
```{r}
dfb = dfa %>% filter(duration>0)

```

creating new columns in data set called duration (length of bike ride), hour_of_day, which is the numeric digit(s) of the started_at cell which is in date/time format. as well as day_of_week
```{r}
dfc= read_excel("DS_bikeshare.xlsx") %>%

  clean_names() %>%
  mutate(duration=as.numeric((ended_at-started_at)/60), hour_of_day=hour(started_at), day_of_week=wday(started_at,label = T))
```


```{r}
dfb=dfa %>% filter(duration>0, )
```

plotting data in terms of duration 
```{r}
dfa %>% ggplot(aes(x=duration)) + geom_histogram(bins=1000) + facet_wrap(vars(weekday))
```

Looking at data plotted in terms of duration as it relates to time of day the bike ride began
```{r}
dfc %>% ggplot(aes(x=started_at, y=duration)) + geom_point()
```

```{r}
dfe=dfa %>% filter(mday(started_at)==3) %>%
slice_sample(n=100)
```
### Tidy data

Separate column term2 into two variables
```{r}
dfa %>% 
  filter(month(started_at)==8,month(ended_at)==8,mday(t)<=14) %>%
  select(start=started_at,ended_at)
```





Creating a dataframe drf, which contains the rideable_type as well as beginning and ending time of ride

```{r}
dfr = dfb %>% 
  filter(month(started_at)==8,month(ended_at)==8) %>%
  select(rideable_type, start=started_at,end=ended_at) %>%
  pivot_longer(start:end,
names_to="type", values_to='t') %>% arrange(t) %>%
  mutate(increment=case_when(
    type=="start"~1,
    type=="end"~ -1
  )) %>%
  group_by(rideable_type) %>%
  mutate(riders=cumsum(increment))
  
```
Creating Dataframe r1 which doesnt include rideable type and doesn't group by rideable type

```{r}
dfr1 = dfb %>% 
  filter(month(started_at)==8,month(ended_at)==8) %>%
  select( start=started_at,end=ended_at) %>%
  pivot_longer(start:end,
names_to="type", values_to='t') %>% arrange(t) %>%
  mutate(increment=case_when(
    type=="start"~1,
    type=="end"~ -1
  )) %>%
  mutate(riders=cumsum(increment))
  
```

How many riders (max) in any hour time interval

```{r}
dfrh=dfr1 %>% group_by(hour(t)) %>% 
  summarise(max_riders=max(riders))
```


```{r}
dfra=dfr1 %>% group_by(hour(t)) %>% 
  summarise(max_riders=max(riders), average_riders=mean(riders))

```


